---
title: "DATASUS - Hospital Admissions (SIH)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Hospital Admissions (SIH)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_hospital_admissions` function provides access to multiple datasets from the **Hospital Information System (SIH)**, which record detailed information about hospital admissions funded by Brazil's public health system (SUS). Each row corresponds to a Hospital Admission Authorization (AIH), and the files are organized by the type of information they contain.
  
***
The `load_hospital_admissions` function offers the following parameters:

1. **dataset**: Specifies the SIH dataset to download:
    * SIH hospitalization data is split across four datasets:
        * `"reduced_aih"` – Reduced AIHs (summary of hospitalizations).
            Contains consolidated information about approved and processed AIHs, including the main procedure performed, related diagnoses, and total costs. This is the most commonly used dataset for statistical and epidemiological analyses.
        * `"professional_services"` – Professional Services performed during hospitalization.
            Provides detailed records of the professional services carried out during hospital stays, including procedures performed, professionals involved (CBO/CNS), and amounts paid for medical and hospital services.
        * `"rejected_aih"` – Rejected AIHs (general reason).
            Includes consolidated records of AIHs that were rejected, specifying the general reason for the rejection but without detailed error codes. Useful for analyzing the volume and impact of rejections.
        * `"rejected_aih_error"` – Rejected AIHs with specific error codes.
            Contains AIHs that were rejected due to inconsistencies found during processing. Each rejection includes a specific error code indicating the reason (e.g., invalid patient data, procedure incompatibilities).

  2. **time_period**: a numeric value or vector indicating the year(s) of the data to be downloaded.
  For example, `2020` or `2015:2020`.
  
  3. **states**: a string or vector of strings indicating the Brazilian state(s) for which the data should be downloaded.
Use `"all"` to download data for the entire country. For specific states (valid only for the `general` dataset), use abbreviations like `"SP"` (São Paulo), `"RJ"` (Rio de Janeiro), or `c("SP", "RJ")`.
  
  4. **raw_data**: Logical, default is `FALSE`.
      * `TRUE`: If TRUE, returns the raw data exactly as provided by DATASUS.
      * `FALSE`: If FALSE (default), returns a cleaned and standardized version of the dataset.

  5. **language**: A string indicating the desired language of variable names and labels. Accepts `"eng"`                      (default) for English or `"pt"` for Portuguese (only when `raw_data = FALSE`).
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download raw data for Reduced AIHs (AIHs Reduzida) – All country, 2010.
data_rd_raw <- load_hospital_admissions(
  dataset = "reduced_aih",
  time_period = 2010,
  states = "AP",
  raw_data = TRUE,
  language = "eng"
)

# Download processed data for Rejected AIHs with Error Codes – State of Amazonas, 2010 to 2020.
# Descriptions in Portuguese.
data_er_processed <- load_hospital_admissions(
  dataset = "rejected_aih_error",
  time_period = 2020,
  states = "AM",
  raw_data = FALSE,
  language = "pt"
)

# Download raw data for Professional Services – States of Rio and São Paulo, 2022.
data_sp_raw <- load_hospital_admissions(
  dataset = "professional_services",
  time_period = 2022,
  states = c("AP", "DF"),
  raw_data = TRUE,
  language = "eng"
)

# Download processed data for Professional Services – Federal District, 2020 to 2022.
# Descriptions in Portuguese.
data_sp_processed <- load_hospital_admissions(
  dataset = "professional_services",
  time_period = 2020:2022,
  states = "DF",
  raw_data = FALSE,
  language = "pt"
)
```
