---
title: "DATASUS - Mortality Data (SIM)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Mortality Data (SIM)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_mortality` function provides access to the **System of Mortality Information (SIM)** datasets, which contain detailed information about deaths in Brazil. Each original SIM data file includes rows corresponding to a declaration of death (DO) and columns with several characteristics of the person, the place of death, and the cause of death.

***
The `load_mortality` function offers the following parameters:
  
  1. **dataset**: Specifies the SIM dataset to download:
      * SIM Datasets:
        * `"general"` – Main Declarations of Death. (National dataset available — `states = "all"`)
            Contains records of all non-fetal Death Certificates (DO) in Brazil, including socio-demographic data, location, and causes of death (ICD-10). It's the base for general mortality analysis.
        * `"fetal"` – Fetal mortality data. (National dataset not available)
        Contains records of fetal deaths, with information on the mother, pregnancy, and causes of fetal death. It's essential for maternal and child health.
        * `"external_causes"` – Mortality data from external causes. (National dataset not available)
            Contains a subset of `"general"` focusing on deaths due to accidents, violence, and other unnatural causes. Used for safety and prevention studies.
        * `"infant"` – Infant mortality data (children). (National dataset not available)
            Contains a subset of `"general"` recording deaths of children under 1 year old, detailing causes and birth-related factors. Crucial for assessing child health.
        * `"maternal"` – Maternal mortality data. (National dataset not available)
            Contains a subset of `"general"` for deaths of women during or shortly after pregnancy/childbirth, detailing obstetric causes. Important for women's health.

  2. **time_period**: a numeric value or vector indicating the year(s) of the data to be downloaded.
  For example, `2020` or `2015:2020`.
  
  3. **states**: (valid only for the `general` dataset) — a string or a vector of strings indicating the Brazilian state(s) for which the data should be downloaded.
The default is `"all"`, which downloads data for the entire country. For specific states, use the official abbreviations such as `"SP"` (São Paulo), `"RJ"` (Rio de Janeiro), or `c("SP", "RJ")`.
  
  4. **raw_data**: Logical, default is `FALSE`.
      * `TRUE`: If TRUE, returns the raw data exactly as provided by DATASUS.
      * `FALSE`: If FALSE (default), returns a cleaned and standardized version of the dataset.
  
  5. **keep_all**: A boolean choosing whether to aggregate the data by municipality, losing individual-level variables (`FALSE`) or to keep all original variables (`TRUE`).
Only applies when `raw_data` is `FALSE`.

  6. **language**: A string indicating the desired language of variable names and labels. Accepts `"eng"`                      (default) for English or `"pt"` for Portuguese (only when `raw_data = FALSE`).
  
  
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download raw data for general mortality - State of Rio de Janeiro, 2022.
raw_data_general_rj <- load_mortality(
  dataset = "general",
  time_period = 2022,
  states = "RJ",
  raw_data = TRUE
)

# Download treated data for general mortality - States of Rio and São Paulo, 2022.
trated_data_general_rj <- load_mortality(
  dataset = "general",
  time_period = 2022,
  states = c("RJ", "SP"),
  raw_data = FALSE,
  keep_all = FALSE # Explicitly stating default behavior
)

# Download treated data for Maternal Deaths - Brazil, 2020 to 2022.
# Descriptions in Portuguese.
# Note: `maternal` does not provide separate files by state.
data_maternal_pt <- load_mortality(
  dataset = "maternal",
  time_period = 2020:2022,
  states = "all",
  raw_data = FALSE,
  language = "pt"
)

# Download treated data for Infant Deaths - Brazil, 2017.
# Keeping all individual variables (not aggregated).
data_infant_full <- load_mortality(
  dataset = "infant",
  time_period = 2017,
  states = "all",
  raw_data = FALSE,
  keep_all = TRUE,
  language = "eng"
)
  
# Download treated data for Fetal Deaths - State of Amazonas, 2000.
data_infant_full <- load_mortality(
  dataset = "fetal",
  time_period = 2000,
  states = "AM", 
  raw_data = FALSE,
  language = "eng"
)

# Download treated data for External Causes Deaths - State of Acre, 2022.
data_infant_full <- load_mortality(
  dataset = "fetal",
  time_period = 2022,
  states = "AC", 
  raw_data = FALSE,
  language = "eng"
)
```
