---
title: "DATASUS - Oncology Cases (Oncology Panel)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Oncology Cases (Oncology Panel)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_oncology_case` function downloads and organizes data from the **Oncology Panel (Painel de Oncologia)**, part of DATASUS. This dataset is widely used in public health and epidemiological analyses related to cancer cases in Brazil.

***
The `load_oncology_case` function offers the following parameters:

  1. **time_period**: a numeric value or vector indicating the year(s) of the data to be downloaded.
  For example, `2020` or `2015:2020`.
  
  2. **raw_data**: Logical, default is `FALSE`.
      * `TRUE`: If TRUE, returns the raw data exactly as provided by DATASUS.
      * `FALSE`: If FALSE (default), returns a cleaned and standardized version of the dataset.

  3. **language**: A string indicating the desired language of variable names and labels. Accepts `"eng"`                      (default) for English or `"pt"` for Portuguese (only when `raw_data = FALSE`).
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download processed oncology data for the year 2023.
# This will return data from the Oncology Panel for all Brazilian states.
oncology_cases_treated <- load_oncology_case(
  time_period = 2023,
  raw_data = FALSE,
  language = "eng"
)

# Download raw oncology data for the years 2021 to 2022 with labels in portuguese.
oncology_cases_raw <- load_oncology_case(
  time_period = 2021:2022,
  raw_data = TRUE,
  language = "pt"
)
```
