---
title: "DATASUS - Outpatient Procedures (SIASUS)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Outpatient Procedures (SIASUS)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_outpatient_procedures` function provides access to various **SIASUS (Ambulatory Information System)** datasets, covering a broad spectrum of outpatient services funded by the public health system (SUS). Each row in these datasets corresponds to a procedure performed at an outpatient level, including clinical, administrative, and financial details. The data is organized by type of service or procedure group.

> **Note:** In all SIASUS datasets, variables related to the *Cadastro Nacional de Saúde* (CNS – National Health Card number) are **encrypted** by DATASUS.  
> This ensures patient confidentiality and means that individual-level CNS identifiers cannot be directly used for linkage across datasets.
> Because of this, this variable is removed when `raw_data = FALSE`.

***
The `load_outpacient_procedures` function offers the following parameters:

  1. **dataset**: Specifies the SIASUS dataset to download:
        * `"ambulatory_production"` – Consolidated Outpatient Procedures (Procedimentos Ambulatoriais).
            Contains records of approved outpatient procedures across all specialties. This is the most comprehensive SIASUS dataset and is often used for general outpatient service analysis. (since Jul/1994 to present)
        * `"bariatric_surgery"` – Pre-Bariatric Surgery (Pré Cirurgia Bariátrica).
            Records related to bariatric surgery procedures performed in outpatient settings. (Jan/2008 to Mar/2013)
        * `"bariatric_surgery_follow_up"` – Bariatric Surgery Follow-Up (Acompanhamento Bariátrico).
            Includes follow-up care for patients who have undergone bariatric surgery, focusing on long-term monitoring and outcomes. (since Apr/2013 to present)
        * `"fistula_confection"` – Vascular Access for Dialysis (Fístula Arteriovenosa).
            Documents procedures involving the creation or maintenance of arteriovenous fistulas, essential for hemodialysis treatment. (since Jun/2014 to present)
        * `"diverse_reports"` – Miscellaneous Specialized Procedures (Laudos Diversos)
            Covers less frequent or highly specialized outpatient procedures not classified in other datasets. (since Jan/2008 to present)
        * `"medicines"` – High-Cost Medications (Medicamentos)
            Tracks the distribution and usage of outpatient medications that are high-cost and part of specific therapeutic programs. (since Jan/2008 to present)
        * `"nephrology"` – Nephrology / Dialysis (Nefrologia)
            Contains outpatient nephrology procedures, particularly related to the care and monitoring of patients with chronic kidney disease. (Jan/2008 to Out/2024)
        * `"dialytic_treatment"` – Dialysis Treatment (Tratamento Dialítico)
            Includes outpatient dialysis treatment sessions for patients with kidney failure. (since Jun/2014 to present)
        * `"psychosocial"` – RAAS Psychosocial Care (RAAS Psicossocial)
            Part of the Specialized Outpatient Mental Health Services. Records care provided through Psychosocial Care Centers (CAPS), including treatments for severe mental disorders and substance use. (since Jan/2013 to present)
        * `"home_care"` – RAAS Home Care (RAAS Atenção Domiciliar)
            Focuses on outpatient care provided at patients' homes, often involving chronic condition management, palliative care, and multi-professional follow-ups. (since Nov/2012 to present)

  2. **time_period**: a numeric value or vector indicating the year(s) of the data to be downloaded.
  For example, `2020` or `2015:2020`.
  
  3. **states**: a string or vector of strings indicating the Brazilian state(s) for which the data should be downloaded.
Use `"all"` to download data for the entire country. For specific states (valid only for the `general` dataset), use abbreviations like `"SP"` (São Paulo), `"RJ"` (Rio de Janeiro), or `c("SP", "RJ")`.
  
  4. **raw_data**: Logical, default is `FALSE`.
      * `TRUE`: If TRUE, returns the raw data exactly as provided by DATASUS.
      * `FALSE`: If FALSE (default), returns a cleaned and standardized version of the dataset.

  5. **language**: A string indicating the desired language of variable names and labels. Accepts `"eng"`                      (default) for English or `"pt"` for Portuguese (only when `raw_data = FALSE`).
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download processed data for Post-Bariatric Surgery Follow-Up (ABO) – State of Acre, 2012.
bariatric_surgery_follow_up <- load_outpatient_procedures(
  dataset = "bariatric_surgery_follow_up",
  time_period = 2012,
  states = "AC",
  raw_data = FALSE,
  language = "eng"
)

# Download processed data for Consolidated Outpatient Procedures (PA) – State of Acre, 2022.
# Descriptions in Portuguese.
ambulatory_production <- load_outpatient_procedures(
  dataset = "ambulatory_production",
  time_period = 2022,
  states = "AC",
  raw_data = FALSE,
  language = "pt"
)

# Download raw data for High-Cost Medications (AM) - State of Pernambuco, 2021.
medicines_raw <- load_outpatient_procedures(
  dataset = "medicines",
  time_period = 2021,
  states = "PE",
  raw_data = TRUE,
  language = "eng"
)

# Download processed data for Psychosocial Care (PS) - State of Acre, 2022 to 2023.
psychosocial <- load_outpatient_procedures(
  dataset = "psychosocial",
  time_period = 2022:2023,
  states = "AC",
  raw_data = FALSE,
  language = "eng"
)
```
