---
title: "DATASUS - Vaccination Data (SI-PNI)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Vaccination Data (SI-PNI)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_vaccines` function provides access to the **National Immunization Program Information System (SI-PNI)**. This dataset contains records of vaccine doses applied across Brazil, allowing for the analysis of immunization coverage and public health strategies.

Unlike other datasets, this function performs automated web scraping on the SI-PNI portal to retrieve the most up-to-date consolidated data directly from the official source.

***
The `load_vaccines` function offers the following parameters:
  
  1. **year**: A numeric value indicating the year of the data to be downloaded.
      * Currently supported range: `1994` to `2022`.
      * Note: You must input only one year at a time.

  2. **state**: A string indicating the Brazilian state abbreviation for which the data should be downloaded (e.g., `"SP"`, `"RJ"`, `"AC"`).
      * Note: You must input only one state at a time.
  
  3. **strategy**: Specifies the vaccination strategy.
      * Common strategies include:
          * `"Rotina"` – Routine vaccination schedule.
          * `"Especial"` – Special immunobiologicals.
          * `"Bloqueio"` – Blocking vaccination in outbreak areas.
          * `"Intensificação"` – Intensification campaigns.
          * `"Serviço Privado"` – Data from private clinics.
      * Note: If set to `NULL`, a selection menu will appear.

  
  4. **product**: Specifies the specific vaccine or immunobiological product (e.g., `"BCG - BCG"`, `"Febre amarela - FA"`, `"Hepatite B - HB"`).
      * The valid options for product depend strictly on the chosen **strategy**.
      * Note: If set to `NULL`, a selection menu will appear.

  5. **language**: A string indicating the desired language of variable names and labels.
      * Accepts `"eng"` (default) for English or ``"pt"`` for Portuguese.
  
  
***
**Interactive Mode:**

If you are unsure of the exact strings for `strategy` or `product`, you can run the function providing only the `year` and `state`. The function will provide an interactive menu in the R console for you to choose from valid combinations.



***
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download data for Yellow Fever (Routine strategy) - State of Acre, 2020.
# Variable names and labels in English.
data_fa_acre <- load_vaccines(
  year = 2020,
  state = "AC",
  strategy = "rotina",
  product = "Febre amarela - FA",
  language = "eng"
)

# Download data for BCG (Routine strategy) - State of São Paulo, 2018.
# Descriptions in Portuguese.
data_bcg_sp <- load_vaccines(
  year = 2018,
  state = "SP",
  strategy = "rotina",
  product = "BCG - BCG",
  language = "pt"
)

# Download data for Private Service strategy - Rio de Janeiro, 2021.
# Specifically for the Hexavalent vaccine.
data_hexa_private <- load_vaccines(
  year = 2021,
  state = "RJ",
  strategy = "serviço privado",
  product = "Hexavalente - HEXA"
)

# Example of calling the function to trigger interactive selection:
# Run this in your console to see the menus.
# data_interactive <- load_vaccines(year = 2022, state = "MG")
```

***
**Technical Note:** As this function relies on web scraping `(chromote)`, ensure you have a stable internet connection and a Chrome-based browser installed on your system.


**Important:** Please be aware that the **SI-PNI** website (*https://sipni.datasus.gov.br/si-pni-web/faces/relatorio/consolidado/dosesAplicadasMensal.jsf*) often experiences significant instability. This may result in connection timeouts, slow response times, or unexpected errors during the scraping process. If the function fails, it is recommended to wait a few minutes and try again. If the error persists, please check the **SI-PNI** portal status or report the issue on our GitHub repository.
