---
title: "DATASUS - Vaccination Data (SI-PNI)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Vaccination Data (SI-PNI)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_vaccines` function provides access to the **National Immunization Program Information System (SI-PNI)**. This dataset contains records of vaccine doses applied across Brazil, allowing for the analysis of immunization coverage and public health strategies.

The function supports **two distinct data ingestion modes**, depending on the availability and stability of the official SI-PNI portals:

  1. **Automated web scraping** of the legacy SI-PNI Web system (recommended for data up to 2022).
  2. **Manual data ingestion**, where the user downloads the data directly from the official DATASUS dashboard and the function performs only cleaning, harmonization, and standardization (required for 2023 onwards).

Both modes produce a **fully harmonized output**, consistent with the historical SI-PNI data structure.

***
The `load_vaccines` function offers the following parameters:
  
  1. **year**: A numeric value indicating the year of the data to be downloaded.
      * Currently supported range: `1994` to `2022`.
      * Note: You must input only one year at a time.

  2. **state**: A string indicating the Brazilian state abbreviation for which the data should be downloaded (e.g., `"SP"`, `"RJ"`, `"AC"`).
      * Note: You must input only one state at a time.
  
  3. **strategy**: Specifies the vaccination strategy.
      * Common strategies include:
          * `"Rotina"` – Routine vaccination schedule.
          * `"Especial"` – Special immunobiologicals.
          * `"Bloqueio"` – Blocking vaccination in outbreak areas.
          * `"Intensificação"` – Intensification campaigns.
          * `"Serviço Privado"` – Data from private clinics.
      * Note: If set to `NULL`, a selection menu will appear.
  
  4. **product**: Specifies the specific vaccine or immunobiological product (e.g., `"BCG - BCG"`, `"Febre amarela - FA"`, `"Hepatite B - HB"`).
      * The valid options for product depend strictly on the chosen **strategy**.
      * Note: If set to `NULL`, a selection menu will appear.
      
  5. **data**: Optional path to a local `.xlsx` file downloaded manually from the DATASUS dashboard.
      * If provided, `load_vaccines()` will **skip web scraping** and only perform data cleaning and harmonization.
      * This argument is **mandatory for data from 2023 onwards**.

  6. **language**: A string indicating the desired language of variable names and labels.
      * Accepts `"eng"` (default) for English or ``"pt"`` for Portuguese.
  
  
***
## Data ingestion modes
### 1. Automated web scraping (SI-PNI Web)

For historical data (`1994`–`2022`), `load_vaccines()` can automatically retrieve consolidated vaccination data directly from the legacy SI-PNI Web portal using web scraping techniques.

This mode:

  1. Requires a stable internet connection.
  2. Depends on the availability and responsiveness of the SI-PNI Web system.
  3. Uses a Chrome-based browser via the `chromote` package.

No manual intervention is required from the user.


### 2. Manual download and harmonization (DATASUS dashboard)

From **2023 onwards**, vaccination data are published exclusively through the new DATASUS interactive dashboard. Due to technical and legal constraints, automated scraping is not supported for this platform.

In this case, the user must **manually download the data**, and `load_vaccines()` will be responsible only for:

  1. Validating the input file
  2. Cleaning column names and labels
  3. Harmonizing dose categories
  4. Standardizing the output to match the historical SI-PNI structure

This approach ensures continuity and comparability with previous years.

#### Step-by-step instructions for manual download

1. Access the DATASUS vaccination dashboard:  
   *https://infoms.saude.gov.br/extensions/SEIDIGI_DEMAS_VACINACAO_CALENDARIO_NACIONAL_OCORRENCIA/SEIDIGI_DEMAS_VACINACAO_CALENDARIO_NACIONAL_OCORRENCIA.html*

2. In the filter panel, **fill only the following fields**:
  * **UF Ocorrência**
  * **Ano Vacina**
  * **Estratégia de Vacinação**
  * **Imunobiológico**
  * **Doses**

   (Do not apply any additional filters)

3. Switch to the **“Tabelas”** tab.

4. In the table configuration:

  * Add the variable **“Tipo Dose”** (at the left side) to the table.
  * Expand all rows to display the full table.

5. Click **“Baixar Dados”** and save the file in `.xlsx` format.

6. Provide the downloaded file to `load_vaccines()` using the `file` argument.


***
**Interactive Mode:**

If you are unsure of the exact strings for `strategy` or `product`, you can run the function providing only the `year` and `state`. The function will provide an interactive menu in the R console for you to choose from valid combinations. (The interactive mode is only valid for data between `1994` and `2022`)



***
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download data for Yellow Fever (Routine strategy) - State of Acre, 2020.
# Variable names and labels in English.
data_fa_acre <- load_vaccines(
  year = 2020,
  state = "AC",
  strategy = "Rotina",
  product = "Febre amarela - FA",
  language = "eng"
)

# Download data for BCG (Routine strategy) - State of São Paulo, 2018.
# Descriptions in Portuguese.
data_bcg_sp <- load_vaccines(
  year = 2018,
  state = "SP",
  strategy = "Rotina",
  product = "BCG - BCG",
  language = "pt"
)

# Download data for Private Service strategy - Rio de Janeiro, 2021.
# Specifically for the Hexavalent vaccine.
data_hexa_private <- load_vaccines(
  year = 2021,
  state = "RJ",
  strategy = "Serviço Privado",
  product = "Hexavalente - HEXA"
)

# Load vaccination data for 2023 using a manually downloaded file
data_fa_acre_2023 <- load_vaccines(
  file = "C:/path/to/downloaded_file.xlsx",
  language = "eng"
)

# Example of calling the function to trigger interactive selection:
# Run this in your console to see the menus.
# data_interactive <- load_vaccines(year = 2022, state = "MG")
```

***
**Technical Note:**

  1. Automated web scraping relies on the legacy SI-PNI Web system and is subject to instability, timeouts, and unexpected failures.
  2. For recent years (2023 onwards), manual download via the DATASUS dashboard is the only supported data source.
  3. Regardless of the ingestion mode, `load_vaccines()` always returns a harmonized dataset with consistent variable names, dose categories, and structure.

**Important:** For the web scraping mode, please be aware that the **SI-PNI** website (*https://sipni.datasus.gov.br/si-pni-web/faces/relatorio/consolidado/dosesAplicadasMensal.jsf*) often experiences significant instability. This may result in connection timeouts, slow response times, or unexpected errors during the scraping process. If the function fails, it is recommended to wait a few minutes and try again. If the error persists, please check the **SI-PNI** portal status or report the issue on our GitHub repository.
