---
title: "DATASUS - Live Births (SINASC)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Live Births (SINASC)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_births` function provides access to the **Live Birth Information System (SINASC)** dataset, which collects and records detailed information about births in Brazil. This data is extracted from Live Birth Certificates (DNVs) and includes information about the newborn, such as sex, weight, and gestational age, as well as data about the mother, such as age, number of children, and health conditions. SINASC is essential for monitoring maternal and child health and generating relevant indicators for public health policy formulation.

***
The `load_births` function offers the following parameters:

  1. **time_period**: A numeric value or vector indicating the year(s) of the data to be downloaded. For      
                      example, 2020 or 2015:2020.
  
  2. **states**: A string or array of strings indicating the Brazilian state(s) for which data should be    
                 downloaded. Use "all" (by default) to download data for the entire country. For specific 
                 states, use abbreviations such as "SP", "RJ", or c("SP", "RJ").
  
  3. **raw_data**: Logical, default is `FALSE`. 
      * `TRUE`: If TRUE, returns the raw data exactly as provided by DATASUS.
      * `FALSE`: If FALSE (default), returns a cleaned and standardized version of the dataset.
  
  4. **language**: A string indicating the desired language of variable names and labels. Accepts "eng"
                   (default) for English or "pt" for Portuguese.
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download raw birth data for 2023 in the state of Rio de Janeiro (RJ).
data_raw_births <- load_births(
  time_period = 2023,
  states = "RJ"
)

# Download raw birth data for 2020 in the states of Rio de Janeiro (RJ) and São Paulo (SP),
# keeping the original raw format.
data_raw_births2 <- load_births(
  time_period = 2020,
  states = c("RJ","SP"),
  raw_data = TRUE
)

# Download raw birth data for 2014 in the state of Amazonas (AM),
# with variable labels in Portuguese.
data_raw_births3 <- load_births(
  time_period = 2014,
  states = "AM",
  language = "pt"
)

# Download processed birth data for 2015 in the state of Amazonas (AM),
# with variable labels in Portuguese for easier analysis.
data_processed_births <- load_births(
  time_period = 2015,
  states = "AM",
  raw_data = FALSE,
  language = "pt"
)
```
