---
title: "DATASUS - Live Births (SINASC)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Live Births (SINASC)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_births` function provides access to the **Live Birth Information System (SINASC)** dataset, which collects and records detailed information about births in Brazil. This data is extracted from Live Birth Certificates (DNVs) and includes information about the newborn, such as sex, weight, and gestational age, as well as data about the mother, such as age, number of children, and health conditions. SINASC is essential for monitoring maternal and child health and generating relevant indicators for public health policy formulation.

***
The `load_births` function offers the following parameters:

  1. **time_period**: A numeric value or vector indicating the year(s) of the data to be downloaded. For      
                      example, 2020 or 2015:2020.
  
  2. **states**: A string or vector of strings indicating the Brazilian state(s) for which the data should be
                 downloaded. Use "all" to download data for the entire country. For specific states, use       
                 abbreviations like "SP", "RJ", or c("SP", "RJ").
  
  3. **raw_data**: Logical, default is `FALSE`. 
      * `TRUE`: if TRUE, returns the raw data exactly as provided by DATASUS.
      * `FALSE`: if FALSE (default), returns a cleaned and standardized version of the dataset.
  
  4. **language**: A string indicating the desired language of variable names and labels. Accepts "eng"
                   (default) for English or "pt" for Portuguese.
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download raw data for the year 2015 in the state of Amazonas (AM).
data_raw_births <- load_births(time_period = 2023,
                               states = "RJ")

data_raw_births <- load_births(time_period = 2020,
                               states = c("RJ","SP"),
                               raw_data = TRUE)


# Download processed data for the year 2015 in the state of Amazonas (AM),
# with variable labels in portuguese.
data_processed_births <- load_births(
  time_period = 2015,
  states = "AM",
  raw_data = FALSE,
  language = "pt"
)
```
