---
title: "DATASUS - Oncology Cases (Oncology Panel)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Oncology Cases (Oncology Panel)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_oncology_case` function downloads and organizes data from the **Oncology Panel (Painel de Oncologia)**, part of DATASUS. This dataset is widely used in public health and epidemiological analyses related to cancer cases in Brazil.

***
The `load_oncology_case` function offers the following parameters:

  1. **raw_data**: there are two options:
      * `TRUE`: If you want the original DATASUS files in their segmented form.
      * `FALSE`: if you want the treated version of the data.
  
  2. **time_period**: Specifies the years for which the data will be downloaded.

  3. **language**: You can choose between Portuguese `("pt")` and English `("eng")` for variable descriptions when `raw_data = FALSE`.
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download processed oncology data for the year 2023.
# This will return data from the Oncology Panel for all Brazilian states.
oncology_data_2023 <- load_oncology_case(
  time_period = 2023,
  raw_data = FALSE,
  language = "eng"
)

# Download raw oncology data for the years 2020 to 2022 with labels in portuguese.
oncology_data_raw <- load_oncology_case(
  time_period = 2020:2022,
  raw_data = TRUE,
  language = "pt"
)
```
