---
title: "DATASUS - Hospital Beds (CNES-LT)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Hospital Beds (CNES-LT)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_hospital_beds` function specifically focuses on the **CNES - LT (Beds)** dataset, part of the National Register of Health Establishments (CNES). This dataset provides information on the number of available hospital beds in health establishments across Brazil.

***
The `load_hospital_beds` function offers the following parameters:

  1. **raw_data**: there are two options:
      * `TRUE`: If you want the original DATASUS files in their segmented form.
      * `FALSE`: if you want the treated version of the data.
      
  2. **time_period**: Specifies the years for which the data will be downloaded.
  
  3. **states**: A vector of states by which to filter the data.

  4. **language**: You can choose between Portuguese `("pt")` and English `("eng")` for variable descriptions when `raw_data = FALSE`.
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download treated data with the number of available beds in Amazonas (AM) and Pará (PA).
data_beds_full <- load_hospital_beds(
  time_period = 2010,
  states = c("AM", "PA"),
  raw_data = FALSE
)

# Download treated data with the number of available beds in whole country.
data_beds_full <- load_hospital_beds(
  time_period = 2010,
  states = "all",
  raw_data = FALSE
)

# Download raw data for the number of hospital beds in 2015 for Rio de Janeiro.
data_beds_raw <- load_hospital_beds(
  time_period = 2015,
  states = "RJ",
  raw_data = TRUE
)
```
