---
title: "DATASUS - Outpatient Procedures (SIASUS)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Outpatient Procedures (SIASUS)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_outpatient_procedures` function provides access to various **SIASUS (Ambulatory Information System)** datasets, covering a broad spectrum of outpatient services funded by the public health system (SUS). Each row in these datasets corresponds to a procedure performed at an outpatient level, including clinical, administrative, and financial details. The data is organized by type of service or procedure group.

***
The `load_outpacient_procedures` function offers the following parameters:

  1. **dataset**: Specifies the SIASUS dataset to download:
    * SIASUS Ambulatory Care Datasets:
        * `"siasus_pa"` – Consolidated Outpatient Procedures (Procedimentos Ambulatoriais).
            Contains records of approved outpatient procedures across all specialties. This is the most comprehensive SIASUS dataset and is often used for general outpatient service analysis.
        * `"siasus_ab"` – Bariatric Surgery (Cirurgia Bariátrica).
            Records related to bariatric surgery procedures performed in outpatient settings.
        * `"siasus_abo"` – Post-Bariatric Surgery Follow-Up (Acompanhamento Bariátrico).
            Includes follow-up care for patients who have undergone bariatric surgery, focusing on long-term monitoring and outcomes.
        * `"siasus_acf"` – Vascular Access for Dialysis (Fístula Arteriovenosa).
            Documents procedures involving the creation or maintenance of arteriovenous fistulas, essential for hemodialysis treatment.
        * `"siasus_ad"` – Miscellaneous Specialized Procedures (Laudos Diversos)
            Covers less frequent or highly specialized outpatient procedures not classified in other datasets.
        * `"siasus_am"` – High-Cost Medications (Medicamentos)
            Tracks the distribution and usage of outpatient medications that are high-cost and part of specific therapeutic programs.
        * `"siasus_an"` – Nephrology / Dialysis (Nefrologia)
            Contains outpatient nephrology procedures, particularly related to the care and monitoring of patients with chronic kidney disease.
        * `"siasus_atd"` – Dialysis Treatment (Tratamento Dialítico)
            Includes outpatient dialysis treatment sessions for patients with kidney failure.
        * `"siasus_ps"` – RAAS Psychosocial Care (RAAS Psicossocial)
            Part of the Specialized Outpatient Mental Health Services. Records care provided through Psychosocial Care Centers (CAPS), including treatments for severe mental disorders and substance use.
        * `"siasus_sad"` – RAAS Home Care (RAAS Atenção Domiciliar)
            Focuses on outpatient care provided at patients' homes, often involving chronic condition management, palliative care, and multi-professional follow-ups.

  2. **time_period**: a numeric value or vector indicating the year(s) of the data to be downloaded.
  For example, `2020` or `2015:2020`.
  
  3. **states**: a string or vector of strings indicating the Brazilian state(s) for which the data should be downloaded.
Use `"all"` to download data for the entire country. For specific states (valid only for the `general` dataset), use abbreviations like `"SP"` (São Paulo), `"RJ"` (Rio de Janeiro), or `c("SP", "RJ")`.
  
  4. **raw_data**: Logical, default is `FALSE`.
      * `TRUE`: If TRUE, returns the raw data exactly as provided by DATASUS.
      * `FALSE`: If FALSE (default), returns a cleaned and standardized version of the dataset.

  5. **language**: A string indicating the desired language of variable names and labels. Accepts `"eng"`                      (default) for English or `"pt"` for Portuguese (only when `raw_data = FALSE`).
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download processed data for Post-Bariatric Surgery Follow-Up (ABO) – State of Acre, 2012.
teste_abo <- load_outpatient_procedures(
  dataset = "siasus_abo",
  time_period = 2012,
  states = "AC",
  raw_data = FALSE,
  language = "eng"
)

# Download processed data for Consolidated Outpatient Procedures (PA) – State of Acre, 2022.
# Descriptions in Portuguese.
teste_pa <- load_outpatient_procedures(
  dataset = "siasus_pa",
  time_period = 2022,
  states = "AC",
  raw_data = FALSE,
  language = "pt"
)

# Download raw data for High-Cost Medications (AM) - State of Pernambuco, 2021.
teste_am_raw <- load_outpatient_procedures(
  dataset = "siasus_am",
  time_period = 2021,
  states = "PE",
  raw_data = TRUE,
  language = "eng"
)

# Download processed data for Psychosocial Care (PS) - State of Acre, 2022.
teste_ps <- load_outpatient_procedures(
  dataset = "siasus_ps",
  time_period = 2022,
  states = "AC",
  raw_data = FALSE,
  language = "eng"
)
```
