---
title: "DATASUS - Hospital Beds (CNES-LT)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Hospital Beds (CNES-LT)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_hospital_beds` function specifically focuses on the **CNES - LT (Beds)** dataset, part of the National Register of Health Establishments (CNES). This dataset provides information on the number of available hospital beds in health establishments across Brazil.

***
The `load_hospital_beds` function offers the following parameters:

  1. **time_period**: a numeric value or vector indicating the year(s) of the data to be downloaded.
  For example, `2020` or `2015:2020`.
  
  2. **states**: a string or vector of strings indicating the Brazilian state(s) for which the data should be downloaded.
Use `"all"` to download data for the entire country. For specific states (valid only for the `general` dataset), use abbreviations like `"SP"` (São Paulo), `"RJ"` (Rio de Janeiro), or `c("SP", "RJ")`.
  
  3. **raw_data**: Logical, default is `FALSE`.
      * `TRUE`: If TRUE, returns the raw data exactly as provided by DATASUS.
      * `FALSE`: If FALSE (default), returns a cleaned and standardized version of the dataset.
  
  4. **keep_all**: A boolean choosing whether to aggregate the data by municipality, losing individual-level variables (`FALSE`) or to keep all original variables (`TRUE`).
Only applies when `raw_data` is `FALSE`.

  5. **language**: A string indicating the desired language of variable names and labels. Accepts `"eng"`                      (default) for English or `"pt"` for Portuguese (only when `raw_data = FALSE`).
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download treated data - States of Amazonas and Pará, 2010.
data_beds_full <- load_hospital_beds(
  time_period = 2010,
  states = c("AM", "PA"),
  raw_data = FALSE,
  language = "eng"
)

# Download treated data - Brrazil, 2010 to 2022.
# Descriptions in Portuguese.
data_beds_full <- load_hospital_beds(
  time_period = 2010:2022,
  states = "all",
  raw_data = FALSE,
  language = "pt"
)

# Download raw data - States of Rio de Janeiro, 2015.
data_beds_raw <- load_hospital_beds(
  time_period = 2015,
  states = "RJ",
  raw_data = TRUE,
  language = "eng"
)
```
