---
title: "DATASUS - Mortality Data (SIM)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DATASUS - Mortality Data (SIM)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `load_mortality` function provides access to the **System of Mortality Information (SIM)** datasets, which contain detailed information about deaths in Brazil. Each original SIM data file includes rows corresponding to a declaration of death (DO) and columns with several characteristics of the person, the place of death, and the cause of death.

***
The `load_mortality` function offers the following parameters:
  
  1. **dataset**: Specifies the SIM dataset to download:
      * SIM Datasets:
        * `"do"` – Main Declarations of Death.
            Contains records of all non-fetal Death Certificates (DO) in Brazil, including socio-demographic data, location, and causes of death (ICD-10). It's the base for general mortality analysis.
        * `"dofet"` – Fetal mortality data.
            Contains records of fetal deaths, with information on the mother, pregnancy, and causes of fetal death. It's essential for maternal and child health.
        * `"doext"` – Mortality data from external causes.
            Contains a subset of `"do"` focusing on deaths due to accidents, violence, and other unnatural causes. Used for safety and prevention studies.
        * `"doinf"` – Infant mortality data (children).
            Contains a subset of `"do"` recording deaths of children under 1 year old, detailing causes and birth-related factors. Crucial for assessing child health.
        * `"domat"` – Maternal mortality data.
            Contains a subset of `"do"` for deaths of women during or shortly after pregnancy/childbirth, detailing obstetric causes. Important for women's health.

  2. **raw_data**: there are two options:
      * `TRUE`: if you want the data as it is originally.
      * `FALSE`: if you want the treated version of the data. Only effective for SIM-DO and subsets, SIH, and CNES-LT.
  
  3. **time_period**: a numeric value or vector indicating the year(s) of the data to be downloaded.
  For example, `2020` or `2015:2020`.
  
  4. **states**: a string or vector of strings indicating the Brazilian state(s) for which the data should be downloaded.
Use `"all"` to download data for the entire country. For specific states, use abbreviations like `"SP"` (São Paulo), `"RJ"` (Rio de Janeiro), or `c("SP", "RJ")`.

  5. **language**: you can choose between Portuguese `("pt")` and English `("eng")`
  
  6. **keep_all**: A boolean choosing whether to aggregate the data by municipality, losing individual-level variables (`FALSE`) or to keep all original variables (`TRUE`).
Only applies when `raw_data` is `FALSE`.
  
**Examples:**

```{r, eval = FALSE}
library(datazoom.saude)

# Download raw data for the year 2022 in the state of RJ for general mortality.
data_sim_do_raw_rj <- load_mortality(
  dataset = "do",
  time_period = 2022,
  states = "RJ",
  raw_data = TRUE
)

# Download treated data with the number of deaths by cause in RJ, aggregated by municipality and year.
data_sim_do_treated_agg_rj <- load_mortality(
  dataset = "do",
  time_period = 2022,
  states = "RJ",
  raw_data = FALSE,
  keep_all = FALSE # Explicitly stating default behavior
)

# Download treated data for Maternal Deaths (SIM-DOMAT) for 2020,
# for the entire country, with descriptions in Portuguese.
# Note: SIM-DOMAT does not provide separate files by state.
data_sim_domat_pt <- load_mortality(
  dataset = "domat",
  time_period = 2020,
  raw_data = FALSE,
  language = "pt"
)

# Download treated data for Infant Deaths (SIM-DOINF) for 2017,
# for all states, keeping all individual variables (not aggregated).
data_sim_doinf_full <- load_mortality(
  dataset = "doinf",
  time_period = 2017,
  raw_data = FALSE,
  keep_all = TRUE
)
```
